---
title: "Nassau and Onongada Plots"
author: "Hope Johnson and Hannah Wheelan"
date: "8/8/2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(lubridate)
library(magrittr)
path <- file.path("/Users","hopecj", "projects", "NY")

```

## Read and munge data

Notice the two parameters in the munge function: "alt_dvoteshare" and "filter_zero". Turning these options on will change results significantly. As-is, we calculate Democratic voteshare as $D votes = D votes \div Total Votes$. Alternatively, one could calculate $D votes = \frac{D votes}{(D votes + R votes)}$. Filtering out district-years where there were zero votes for a Democrat or Republican candidate was not done for either counties. 

```{r cleaner data}
# this function requires the following columns:
#  "Year", "District", "D.Voteshare"

# default D_voteshare for Onondaga cnty is D.voteshare = Dvotes/total votes


# alt_dvoteshare is binary, if "TRUE" then D.voteshare = D/(D + R)
# filter_zero is binary, removes uncontested races (do this thoughtfully!)
# redistrict_start is a date like "2011", and sets the variable "pre_redistricting" to 1 BEFORE (not incl.) that date
# impute_zero assumes a fraction of some number for uncontested races (default is .25)
clean_file <- function(path, alt_dvoteshare = FALSE, filter_zero = FALSE, 
                       redistrict_start, impute_zero = .25) {
  dat <- read.csv(path)
  if (filter_zero) {
    dat %<>%
      filter(Dem.Votes != 0,
             GOP.Votes != 0)
  }
  if (alt_dvoteshare) {
    dat$D.Voteshare <- dat$D.Voteshare2
  }
  dat$Year <- lubridate::ymd(dat$Year, truncated = 2L)
  # create indicator for pre/post-redistricting
  pre_redist <- lubridate::ymd(redistrict_start, truncated = 2L) - years(1)
  pre_redist_span <- c(seq(min(dat$Year), pre_redist, by = "years"))
  dat$pre_redistricting <- ifelse(dat$Year %in% pre_redist_span, 1, 0)
  
  # for races where there was no democratic candidate, give Dem voteshare `impute_zero` 
  # for races wher there was no gop candidate, give Dem voteshare (1 - `impute_zero`)
  # 25% is default, 0% means no change from actual results
  dat$D.Voteshare <- ifelse(dat$D.Voteshare == 0, impute_zero, 
                            ifelse(dat$D.Voteshare == 1, (1 - impute_zero), dat$D.Voteshare))
  
  ds_1 <- do.call(rbind, lapply(split(dat, dat$District), function(d) {
  data.frame(mean = mean(d$D.Voteshare), sd = sd(d$D.Voteshare), District = d$District)
  }))
  ds_2 <- ds_1 %>%
    group_by(District) %>%
    add_tally()
  
  ds <- ds_2 %>%
    mutate(se = sd/sqrt(n))
  
  ds_tomerge <- distinct(ds)
  
  dat %<>%
    left_join(ds_tomerge, by = "District")
  
  return(dat)
}
```

# Nassau County 

```{r}
n_path <- file.path(path, "raw nassau", "Nassau_County_Legislature_Election_Results.csv")
dat <- clean_file(n_path, redistrict_start = "2013")
skim(dat)
```

What is the average two-party voteshare in Nassau County (across all districts)?

```{r}
dat %>%
  group_by(Year) %>%
  skim(Party, D.Voteshare)
```



### Dem Voteshare by District

A Dem voteshare close to 50% before redistricting, and far below or above 50% after redistricting, might point to packing or cracking. 

- District 5!
- District 3
- District 7
- District 19

```{r echo=FALSE, include=FALSE}
# NOT a CONFIDENCE INTERVAL, JUST MEAN ± SD
ggplot() +
  geom_point(data = dat, aes(District, mean), colour = 'grey', size = 2) +
  geom_errorbar(
    data = dat,
    aes(District, mean, ymin = mean - sd, ymax = mean + sd), # NOT a conf. int, 
    colour = 'grey',
    width = 0.4
  ) + 
  geom_point(data = dat, aes(District, D.Voteshare, colour = factor(pre_redistricting))) +
  scale_x_continuous(breaks = 1:20) + 
  labs(y = "D Voteshare")
```

```{r warning=FALSE}
# 95% CONFIDENCE INTERVAL
ggplot() +
  geom_point(data = dat, aes(District, mean), colour = 'grey', size = 2) +
  geom_errorbar(
    data = dat,
    aes(District, mean, ymin = mean - 1.96*se, ymax = mean + 1.96*se), 
    colour = 'grey',
    width = 0.4
  ) + 
  geom_point(data = dat, aes(District, D.Voteshare, colour = factor(pre_redistricting))) +
  scale_x_continuous(breaks = 1:20) + 
  labs(y = "D Voteshare", title = "Nassau County Democratic Voteshare by District") + 
  scale_color_manual(name="Pre-redistricting", 
                        labels = c("No", 
                                   "Yes"), 
                        values = c("0"="lightblue", 
                                   "1"="orange"))

```


### Dem Voteshare Over Time

Re-districting happened in 2011 in Nassau County.

```{r}
break.vec <- seq(from = as.Date("2009-01-01"), to = as.Date("2017-01-01"),
                 by = "2 years")

dat %>%
  filter(se > 0.03026350) %>%
  ggplot() + 
  geom_point(aes(Year, D.Voteshare, colour = factor(District))) + 
  scale_x_date(breaks = break.vec, date_labels = "%Y") + 
  geom_hline(yintercept = .5) + 
  geom_line(aes(Year, D.Voteshare, colour = factor(District)), linetype = 2) + 
  labs(colour="District", 
       title = "Democratic Voteshare Over Time in Nassau County", 
       subtitle = "(showing districts with std. error > 0.03)") 
```


# Onondaga County 

```{r}
# NB: in onongada county, D.Voteshare is calculated by (D Votes/TOTAL VOTES) here
#     
o_path <- file.path(path, "raw onondaga", "Onondaga County Legislature.csv")
o_dat <- clean_file(o_path, redistrict_start = "2011", alt_dvoteshare = TRUE)
skim(o_dat) 
```
What is the average two-party voteshare in Onondaga County (across all districts)?

```{r}
o_dat %>%
  group_by(Year) %>%
  skim(Party, D.Voteshare)
```

A word of caution for the Onondaga results. There are some peculiarities in the data, perhaps due to re-districting (post-2009).

```{r}
o_dat %>% 
  filter(District %in% c(1, 2)) %>% 
  arrange(District) %>%
  select(District, Year, D.Voteshare)
```

### Dem Voteshare by District

```{r warning=FALSE}
# 95% CONFIDENCE INTERVAL
ggplot() +
  geom_point(data = o_dat, aes(District, mean), colour = 'grey', size = 2) +
  geom_errorbar(
    data = o_dat,
    aes(District, mean, ymin = mean - 1.96*se, ymax = mean + 1.96*se), 
    colour = 'grey',
    width = 0.4
  ) + 
  geom_point(data = o_dat, aes(District, D.Voteshare, colour = factor(pre_redistricting))) +
  scale_x_continuous(breaks = 1:19) + 
  labs(y = "D Voteshare", title = "Onondaga County Democratic Voteshare by District") + 
  scale_color_manual(name="Pre-redistricting", 
                        labels = c("No", 
                                   "Yes"), 
                        values = c("0"="lightblue", 
                                   "1"="orange"))


```

### Dem Voteshare Over Time

```{r}
break.vec <- seq(from = as.Date("2009-01-01"), to = as.Date("2017-01-01"),
                 by = "2 years")

o_dat %>%
  filter(se >0.048648) %>%
  ggplot() + 
  geom_point(aes(Year, D.Voteshare, colour = factor(District))) + 
  scale_x_date(breaks = break.vec, date_labels = "%Y") + 
  geom_hline(yintercept = .5) + 
  geom_line(aes(Year, D.Voteshare, colour = factor(District)), linetype = 2) + 
  labs(colour="District", 
       title = "Democratic Voteshare Over Time in Onondaga County", 
       subtitle = "(showing districts with std. error > 0.048)") 
```

